Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-28 paper_jasa.tex
caption changed for loess figure
subecjt-specific variation

@@

It is really curious to see that the power of the visual test is really high when effect is small. These results are based on whether the subject correctly picked the actual data plot, regardless of reason. Although they were asked to select the plot that exhibited the highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data. This will be explored further in Section \ref{sec:TypeIII}.

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
       \caption{Power in comparison to effect for the three experiments. Points indicate subject responses, with size indicating count. Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup. The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.  Conventional test power is drawn as a dashed line. For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
}
       \label{fig:power_loess_effect}
\end{figure}

@@

In all the three experiments, each subject saw a lineup only once. This allows no random effect for lineup and this simplifies model \ref{mixed} as $\tau_{\ell  i}=\tau_i$.   This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}. Different subject performs differently depending on effect as well. So, we include a random interaction term with effect. Parameter estimates of the model is shown in table \ref{tbl:model_par}.  The estimated subject specific power curves are shown in figure \ref{fig:power_mixed_subject}. Subject specific power is quite different for three experiments. It appears that scatter plot serves better job compared to box plot as a visual test statistic. But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.

>>>>

\begin{figure}[hbtp]
   \centering
       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
       \caption{Power in comparison to effect for the three experiments. Points indicate subject responses, with size indicating count. Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup. The loess curve (blue) estimates the observed power, and grey bands show simultaneous bootstrap confidence band.  Conventional test power is drawn as a dashed line. For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data. Power for contaminated data is shown by cross mark.
}
       \label{fig:power_loess_effect}
\end{figure}

@@

It is really curious to see in the results of experiment 3 that the power of the visual test is really high when effect is small. These results are based on whether the subject correctly picked the actual data plot, regardless of reason. Although they were asked to select the plot that exhibited the highest association between the two variables, it may have been that they were cueing to the cluster of contaminated data. This will be explored further in Section \ref{sec:TypeIII}.

@@

@@

\green{Do we have more than one model?? If so then it is appropriate to refer to them as model 1, model 2, ...., but if there is only one model then we need to refer readers back to the equation, occasionally.}

\green{Actually I think this paragraph should have already been covered when the model was introduced. At this point we just need to discuss the results, so I'm going to cut the paragraph and start again.}

%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{mixed}).  For all the three experiments, each subject saw a lineup just once. This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i}=\tau_i$. The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$. So, we include a random interaction term with effect. 
Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.  The estimated subject specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines). The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test. Subject-specific power is quite different for three experiments. In experiment 1 subjects performed similarly, and substantially worse than the conventional test. In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects. In experiment 3 there is the most subject-specific variation. Some subjects performed substantially better than the conventional test, and on average the visual test was better.

\green{The dashed line cannot be seen in experiment 2 for small effect, and it is essential to see this at this point. What can you do to make the dashed line visible?}
